---
title: "eco_imp_water"
author: "Mbye Sallah"
date: "2024-08-07"
output: html_document
---

#### Import libraries
```{r}
library(tidyverse)
library(terra)
library(sf)
library(tigris)
library(raster)
```


#### read the land condition dataset
```{r}
root_dir <- config::get()
land_condition <- rast(paste0(root_dir, "/private/datasets/raw_data/Landscape Condition/Americas_N_LCM_Cat100.tif"))
land_condition
```

#### Get the summary info of the land contition data
```{r}
summary(land_condition)
```



#### Plot the raster data
```{r}
plot(land_condition)
```


#### change the values greater than 100 to NA
```{r}
land_condition[land_condition > 100] <- NA
land_condition
```

#### Check the summary after modifying the values
```{r}
summary(land_condition)
```

#### Plot of the modified data
```{r}
plot(land_condition)
```


#### Invert land condition values by subtracting 100
```{r}
inverted_land_condition <- 100 - land_condition
inverted_land_condition
```


#### Read the water service area shapefile
```{r}
wsa_boundaries <- read_sf(paste0(root_dir, "/private/datasets/raw_data/EPA PWS boundaries/EPA_CWS_V1.shp"))
```



#### Use the crosswalk_spatial function to do the crosswalk
```{r}
source(paste0(root_dir, "/code/crosswalk/crosswalk_spatial.R"))
land_condition_wsa <- crosswalk_spatial(data = inverted_land_condition, 
                                          target = wsa_boundaries,
                                          join_method = "areal_weighted")
```



#### read the bio_imp data
```{r}
bio_imp <- rast(paste0(root_dir, "/private/datasets/raw_data/Biodiversity Importance/mobi_rsr_crf.tif"))
```

#### Get the bio importance data on water service area
```{r}
bio_imp_wsa <- crosswalk_spatial(data = bio_imp, 
                                target = wsa_boundaries,
                                join_method = "areal_weighted")
```

#### Join land condition and bio imp
```{r}
eco_imp_land <- land_condition_wsa %>%
  st_join(bio_imp_wsa, by = "PWSID") %>%
  mutate(eco_imp_land_mean = var_mean.x * var_mean.y,
         eco_imp_land_max = var_max.x * var_max.y) %>%
  dplyr::select(PWSID.x, eco_imp_land_mean, eco_imp_land_max) %>% 
  rename(PWSID = PWSID.x) %>% 
  st_drop_geometry()
```


#### Save the dataset
```{r}
write_csv(eco_imp_land, paste0(root_dir, "/private/datasets/cleaned_data/wsa_eco_imp_land.csv"))
```


