---
title: "Eco Importance Water"
author: "Mbye Sallah"
date: "2024-08-05"
output: html_document
---

#### Import libraries
```{r}
library(tidyverse)
library(terra)
library(sf)
library(tigris)
library(raster)
```



#### Read RSEI_stream (Risk-Screening Environmental Indicators)
```{r}
root_dir <- config::get()
ejscreen_24_cnmi <- read_csv(paste0(root_dir, "/private/datasets/raw_data/EJSCREEN_2024_Tract_with_AS_CNMI_GU_VI.csv"))
```

#### Rename PWDIS to RSEI_stream
```{r}
ejscreen_24_cnmi <- ejscreen_24_cnmi %>%
  rename(RSEI_stream = PWDIS)
```

#### Save the daataset with the renamed column
```{r}
write_csv(ejscreen_24_cnmi, "~/DSSG/WaterReuseDSSG2024/private/datasets/cleaned_data/EJSCREEN_2024_Tract_with_AS_CNMI_GU_VI.csv")
```

### Get census tract shapefile
```{r}
tract_sf <- tracts(cb = TRUE, year = 2023)
crs(tract_sf)
```



#### Left join the ejscreen_24_cmni and attains
```{r}
ejscreen_24_cnmi_tract <- ejscreen_24_cnmi %>%
  dplyr::select(ID, RSEI_stream) %>%
  left_join(tract_sf, join_by(ID == GEOID))
```

#### Read the biodiversity importance dataset
```{r}
bio_imp <- rast("~/DSSG/WaterReuseDSSG2024/private/datasets/raw_data/Biodiversity Importance/mobi_rsr_crf.tif")
bio_imp
```
#### plot the bio_imp
```{r}
plot(bio_imp)
```

#### Check the unit of the bio_imp
```{r}
units(bio_imp)
```




<!-- #### Extract the census tract level data of bio_imp -->
<!-- ```{r} -->
<!-- source(paste0(root_dir, "/code/crosswalk/get_raster_data.R")) -->
<!-- extrated_bio_imp <- get_raster_data(bio_imp, tract_sf, fun = mean) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- write_csv(extrated_bio_imp, paste0(root_dir, "/private/datasets/cleaned_data/tract_aggregated_bio_imp.csv")) -->
<!-- ``` -->



<!-- #### Open raster data using Raster -->
<!-- ```{r} -->
<!-- bio_imp_raster <- raster("~/DSSG/WaterReuseDSSG2024/private/datasets/raw_data/Biodiversity Importance/mobi_rsr_crf.tif") -->
<!-- ``` -->


#### Do the crosswalk using crosswalk_raster function
```{r}
source(paste0(root_dir, "/code/crosswalk/crosswalk_spatial.R"))
bio_imp_interpolated <- crosswalk_spatial(data = bio_imp, 
                 target = tract_sf,
                 join_method = "areal_weighted")
```


<!-- #### Save the interpolated data to csv and shapefile -->
<!-- ```{r} -->
<!-- write_csv(bio_imp_interpolated, paste0(root_dir, "/private/datasets/cleaned_data/bio_imp_tract_interpolated.csv")) -->
<!-- st_write(bio_imp_interpolated, paste0(root_dir, "/private/datasets/cleaned_data/bio_imp_tract_interpolated.shp")) -->
<!-- ``` -->



#### Multiply the census tract bio_imp aggregated data with RSEI_stream
```{r}
bio_imp_data <- bio_imp_interpolated %>% 
  left_join(ejscreen_24_cnmi_tract, join_by(GEOID == ID)) %>% 
  mutate(eco_imp_water_mean = var_mean * RSEI_stream,
         eco_imp_water_max = var_max * RSEI_stream) %>%
  dplyr::select(eco_imp_water_mean, eco_imp_water_max, geometry.x) %>% 
  rename(geometry = geometry.x)
```

#### Read the water service area shapefile
```{r}
wsa_boundaries <- read_sf(paste0(root_dir, "/private/datasets/raw_data/EPA PWS boundaries/EPA_CWS_V1.shp"))
```


#### Do the crosswalk using crosswalk_spatial
```{r}
source(paste0(root_dir, "/code/crosswalk/crosswalk_spatial.R"))
wsa_eco_bio_water <- crosswalk_spatial(data = bio_imp_data,
                                       target = wsa_boundaries,
                                       join_method = "areal_weighted")
```


```{r}
wsa_eco_bio_water <-  wsa_eco_bio_water %>% 
  dplyr::select(PWSID, eco_imp_water_mean, eco_imp_water_max)
```



#### Save the data in csv format
```{r}
write_csv(wsa_eco_bio_water, paste0(root_dir, "/private/datasets/cleaned_data/wsa_eco_bio_water.csv"))
```

