---
title: "Water Reuse Data Cleaning - NPDES DMRs"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

<!-- # Executive Summary -->
<!-- * This document computes `path_load` and `temp_load` variables -->
<!-- * `path_load` has null values `r na_counts_newvar$path_load`% -->

---
```{r}
library(sf)
library(dplyr)
library(conflicted)


conflicts_prefer(dplyr::select)
```

Loading the dataset:

```{r}
datasets_root <- config::get("datasets_path")
dataset_path <- "Drivers/DMR_combined_WW/DMR_combined_WW.csv"
df <- read.csv(paste0(datasets_root, dataset_path))
#View(df)
```

Examining where there are `NA` values:

```{r}
na_counts <- sapply(df, function(x) mean(is.na(x)))
print(na_counts)
```
A tiny number of facility latitudes and longitudes are `NA`. These need to be dropped, otherwise the geospatial functions employed later in this analysis will throw errors.

```{r}
conflicts_prefer(dplyr::filter)
df <- df %>% filter(!is.na(Facility_Latitude) & !is.na(Facility_Longitude))
```
First task: construct `path_load` variable. `path_load = Path_count_max * Path_Avg_flow`.

```{r}
df$path_load = df$Path_count_max * df$Path_Avg_Flow
```

Second task: construct `temp_load` variable. `temp_load = Temp_max * Temp_Avg_Flow`.

```{r}
df$temp_load = df$Temp_max * df$Temp_Avg_Flow
```

Let's check how many `NA` values are in the constructed variables:

```{r}
na_counts_newvar <- df %>% 
  select(c("FRS_ID", "Path_count_max","Path_Avg_Flow",
           "path_load", "Temp_max", "Temp_Avg_Flow", "temp_load")) %>% 
  sapply(function(x) mean(is.na(x)))

print(na_counts_newvar)
```


```{r}
wsa_filename = "EPA PWS Boundaries/EPA_CWS_V1.shp"

wsa_path <- paste0(config::get("datasets_path"), wsa_filename)
wsa <- st_read(wsa_path)

source("code/crosswalk/crosswalk_spatial.R")

mapped_df <- crosswalk_spatial(df, wsa, location_columns = c("Facility_Latitude", "Facility_Longitude"))

out_filename <- "processed_datasets/npdes_dmrs.shp"
out_path <- paste0(config::get("datasets_path"), out_filename)

columns_to_rename <- setdiff(colnames(mapped_df), "geometry")

colnames(mapped_df)[colnames(mapped_df) %in% columns_to_rename] <- paste0("x", seq_along(columns_to_rename))

st_write(mapped_df, out_path)
```

---
A high percentage of values are `NA`. Let's investigate if it would be possible to impute them with MLR.

```{r}
numeric_df <- df[sapply(df, is.numeric)]

# use argument tells R to ignore NAs
cor_matrix <- cor(numeric_df, use="pairwise.complete.obs")

# puts the cor_matrix in long form, making it easier to filter the cross-correlations
melted_cor_matrix = melt(cor_matrix)

# Full correlation matrix is unwieldy, so focus on the high correlations.
threshold <- 0.5

filtered_correlations <- melted_cor_matrix %>%
  filter(abs(value) > threshold & Var1 != Var2) %>% 
  filter(Var1 == "Temp_max" | 
           Var1 == "Temp_Avg_Flow" |
           Var1 == "Path_count_max" |
           Var1 == "Path_Avg_Flow") %>%
  arrange(Var1)

print(filtered_correlations)
```

`Temp_max` is 86% `NA`. Unfortunately, it is highly correlated only with another variable (`Temp_avg`) with a high percentage of `NA` values, so we cannot use linear regression to try to impute it.
---

# Understanding high NAs in Temp_max

Hypothesis: Large facilities (measured by flow volume) are likelier to collect temperature data than smaller facilities. 

Test: bucket facilities into quintiles by flow volume size; calculate percentage of `NA` values in each quintile.

```{r}
# divide into quintiles
df_ <- df %>% 
  mutate(flow_quintile = ntile(Total_Annual_Flow_MMGal, 5)) %>%
  select(c(FRS_ID, Facility_Name, City, State, flow_quintile, Temp_max))

quintile_summary <- df_ %>% 
  group_by(flow_quintile) %>%
  summarize(
    total_count = n(),
    na_count = sum(is.na(Temp_max)),
    na_percentage = (na_count / total_count) * 100
  )

print(quintile_summary)
```
The result indicates that the hypothesis is correct - higher-flow facilities are likelier to measure and report `Temp_max`. However, even in the group with the highest flow (quintile 5 = highest), the large majority of `Temp_max` observations are `NA`.

