---
title: "Crosswalking play file"
output: html_document
date: "2024-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pacman)


p_load(usmap, raster, exactextractr, tidycensus,tidyverse,sf,data.table,tigris,car,janitor,rmapshaper, terra, qgisprocess)

```

## Testing CWNS example (points and counties shapefile)
# Counties in new column of joined dataframe should match existing county column from cwns dataset

```{r}
# Load Tigris County shapefile, CWNS dataset 

options(tigris_use_cache = TRUE)

# Load the counties shapefile for the entire United States
counties_sf <- counties(year = 2023, cb = TRUE)

cwns <- read_csv("~/Downloads/cwns-flow.csv")

#Select only a few columns
cwns <- cwns %>% 
  select(COUNTY_FIPS, LATITUDE, LONGITUDE, CURRENT_DESIGN_FLOW, LOCATION_TYPE)

#Convert cwns to shapefile using counties CRS  
cwns_sf <- st_as_sf(cwns, coords = c("LONGITUDE", "LATITUDE"), crs = st_crs(counties_sf))

```


## Testing Interbasin Water Transfers Example (lines in counties shapefile)
# Lines should be new counties column should be populated in addition to existing state and country columns from ibt dataset
```{r}
# Load ibt data
ibt <- read_sf(dsn = '~/Downloads/IBT Geospatial Data/', layer = "Digitized_IBTs")

# Transform CRS to Counties CRS
ibt <- st_transform(ibt, crs = st_crs(counties_sf))

```

## Point-to-boundary  
```{r}

crosswalk_geom <- function(data, shapefile, location_columns = NULL, type = "shapes") {
  
  
  if (type == "points" && !is.null(location_columns)) {
    # Convert the dataframe to a spatial object using the shapefile's CRS
    points <- st_as_sf(data, coords = location_columns, crs = st_crs(shapefile))
    
    # Perform the spatial join to find the containing shape for each point
    joined <- st_join(points, shapefile, join = st_within, left = TRUE)
  } else {
    # Ensure data is a spatial object and has the same CRS as the shapefile
    data <- st_transform(data, crs = st_crs(shapefile))
    data <- st_make_valid(data)
    
    # Perform the spatial join
    joined <- st_join(data, shapefile, join = st_contains)
  }
  
  return(joined)
}


```

```{r}
crosswalk_geom(cwns, counties_sf, c("LONGITUDE", "LATITUDE"), type = "points")
```
```{r}
#Attempt crosswalk on point data via st_interpolate 
library(stars)

raster_example <- raster("~/Downloads/CONUS_ww_raster.tif")

```

```{r}

crosswalk_raster <- function(data1, data2, method = "area", output_type = NULL, points = FALSE) {
  # Function for Raster/Raster
  
  if (st_crs(data1) != st_crs(data2)) {
    data2 <- st_transform(data2, st_crs(data1))
  }
  
  combine_rasters <- function(raster1, raster2, method) {
    if (method == "area") {
      combined <- (raster1 + raster2) / 2  # Example: averaging the rasters
    } else if (method == "population") {
      combined <- raster1 + raster2  # Example: summing the rasters
    } else {
      stop("Invalid method for raster combination.")
    }
    return(combined)
  }

  # Function for Raster/Point
  raster_point_join <- function(raster, points) {
    point_values <- exact_extract(raster, points, fun = "mean", progress = FALSE)
    point_values <- unlist(point_values)
    point_values[is.na(point_values)] <- 0  # Handle NA values
    combined <- cbind(st_as_sf(points), raster_value = point_values)
    return(combined)
  }

  # Function for Raster/Shapefile
  raster_shapefile_join <- function(raster, shapefile, output_type) {
    if (output_type == "shapefile") {
      extracted_values <- exact_extract(raster, shapefile, "mean")  # Example: mean values
      shapefile$extracted_values <- extracted_values
      return(shapefile)
    } else if (output_type == "raster") {
      shapefile_rasterized <- rasterize(shapefile, raster)
      return(shapefile_rasterized)
    } else {
      stop("Invalid output type for raster/shapefile combination.")
    }
  }

  # Determine the types of input data and call the appropriate function
  if (inherits(data1, "Raster") && inherits(data2, "Raster")) {
    return(combine_rasters(data1, data2, method))
  } else if (inherits(data1, "Raster") && inherits(data2, "sf") && points== TRUE) {
    return(raster_point_join(data1, data2))
  } else if (inherits(data1, "Raster") && inherits(data2, "sf")) {
    return(raster_shapefile_join(data1, data2, output_type))
  } else {
    stop("Unsupported data types.")
  }
}
```

```{r}
combined <- crosswalk_raster(raster_example, counties_sf, output_type = "shapefile")

```

```{r}
head(combined)
```



## Polygon version
```{r}
# Determine which counties the polygons are "in" 

ibt <- st_transform(ibt, crs = st_crs(counties_sf))

try_join <- st_join(ibt, counties_sf, join = st_within)
head(try_join)

```
## Test on watershed to county 

```{r}

aquifers <- st_read("~/Downloads/aquifers_us/us_aquifers.shp")
head(aquifers)

aquifers <- st_transform(aquifers, crs = st_crs(counties_sf))
aquifers <- st_make_valid(aquifers)

joined_sfs <- crosswalk_geom(aquifers, counties_sf)

```



```{r}

plot_data <- usmap::usmap_transform(joined_sfs)

p <- usmap::plot_usmap("states", color = "gray80") + 
      geom_sf(data = plot_data, color = "lightblue", size = 0.2)

print(p)
```


##Test on rasters 

## Notes for convert_to_shapefile code 
```{r}
#general conversion from csv and excel sheet containing latitude and longitude columns to shapefile 
# convert_to_shapefile <- function(data, location_columns, crs){
#   #Args: 
#     # data: a csv or excel spreadsheet containing latitude and longitude columns
#     # 
#   cwns_sf <- st_as_sf(cwns, coords = , crs = st_crs(counties_sf))
# 
# cwns_sf$geometry
# 
# counties_sf$geometry
# }

```
