---
title: "drought risk data procesing"
output: html_document
date: "2024-06-26"
---

## Introduction


```{r download}
library(ipumsr)
library(plyr)
library(dplyr)
data <- read_xlsx("data/final_results_dissemination_NDINDC.xlsx")
```

##Processing
I merged county-level base map with dataset on NDI, NDC information. Merged output is named "data_plot," which I use for mapping.
```{r}
library(urbnmapr)

data$county_fips <- data$FIPS

counties <- get_urbn_map(sf = TRUE, map = "counties")
counties$county_fips <- as.numeric(counties$county_fips)


#resulting object inherits the class of the object on the left (order matters!)
data_plot <- left_join(counties, 
                        data, 
                        by = "county_fips")

str(data_plot)
data_plot$NDC <- as.numeric(data_plot$NDC)
```

## Mapping
The decision to use the 98th percentile as the cutoff for visualization is based on the distribution of NDI values. Upon reviewing the histogram of the NDI column, it became evident that a handful of outliers (ex. Virginia) have NDI values greater than 200, while the majority of counties show NDI values below 1. Excluding these outliers from the visualization enhances clarity in distinguishing between counties.

```{r}
library(ggplot2) 

plot_percentile <- function(data, variable, caption, title,
                            percentile=0.95,
                            low_color = "white", high_color = "red",
                            na_color = "gray"){
  percentiles <- quantile(data[[variable]], probs=c(0, percentile), na.rm=TRUE)
  
  plot_data_variable <- ggplot(data)+
    geom_sf(mapping = aes_string(fill = variable)) +
    scale_fill_gradient(name = variable, 
                        label = scales::comma_format(),
                        low=low_color,
                        high=high_color,
                        na.value = na_color,
                        limits = percentiles,
                        breaks = scales::pretty_breaks(n = 5)) + 
    labs(title = title,
         caption = caption) +
    theme(text = element_text(family = "Times New Roman"))
  
      return(plot_data_variable)
}

plot_percentile(data = data_plot, 
                variable = "NDC", 
                title = "NDC by county", 
                caption = "NDC is the cumulative deficit over 1949-2010.")

plot_percentile(data = data_plot, 
                variable = "NDI", 
                title = "NDI by county", 
                caption = "NDI is the annual cumulative deficit for the worst drought year.")
```

## References
Devineni, N., Lall, U., Etienne, E., Shi, D., & Xi, C. (2015). 
America’s water risk: Current demand and climate variability. Geophysical Research Letters, 1–9. doi:10.1002/2015GL063487.




