---
title: "Salt Water Intrusion"
output: html_document
date: "2024-07-01"
---

This script is intended to map "salt water intrusion" data onto the coastline shapefile.

## Data-specific features
- This involves aggregation process by grouping different COMIDs up to the REACHCODE level.
- Minor thing, but saltwater intrusion is a redflag. I used red to indicate its harmfulness, but might want to think about general red-blue color range when generalizing. 

## General features
- This is a common problem across many datasets, but seems that shape file and substantive data do not always share the same type and name. 

## Introduction

- ComID is about unique feature identifier in NHDPlusV2.
- REACHCODE is a common identifier for Catchment_CONUS_coastline.shp file and Saltwater Intrusion dataset.

```{r download}
library(ipumsr)
library(plyr)
library(dplyr)
library(readr)
library(sf)
library(xml2)

data <- read_csv("~/Desktop/WaterReuseDSSG2024/data/SGD_Coastal_Vulnerabilities.csv", 
                 col_types = cols(.default = "c"))
map <- read_sf("~/Downloads/Catchment_CONUS_coastline.shp")


data$REACHCODE <- ifelse(nchar(data$REACHCODE) < 14, 
                          paste0("0", data$REACHCODE), 
                          data$REACHCODE)
 
# map$REACHCODE <- ifelse(nchar(map$REACHCODE) < 14, 
#                          paste0("0", map$REACHCODE), 
#                          map$REACHCODE)


#I'm aggregating ComID-level observations up to REACHCODE level. This is because "map" is at the REACHCODE level.

data <- data %>%
  group_by(REACHCODE, REGION) %>%
  summarize(SWIVULN = sum(as.numeric(SWIVULN)))

#sum(map$REACHCODE%in% data$REACHCODE)
#sum(data$REACHCODE%in% map$REACHCODE)

#hold <- map[map$REACHCODE %in% data$REACHCODE[-which(data$REGION %in% c("TX", "CA","PN"))],]

usmap <- usmap_transform(map)

plot_data <- usmap %>%
  left_join(data, by = c("REACHCODE"))

plot_usmap("counties", color = "gray80") + 
  geom_sf(data = plot_data, aes(fill = SWIVULN))



plot_usmap("counties", color = "gray80") + 
  geom_sf(data = plot_data, color = "lightblue", size = 0.2) +
  geom_point(data = plot_data,
    aes(color = SWIVULN, size = SWIVULN, geometry = geometry),
    alpha = 0.2,
    color = "red",
    stat = "sf_coordinates"
  ) +
  scale_size_continuous(name = "Saltwater Vulnerability",
                        labels = scales::comma_format(),
                        range = c(0.01, 7)) + 
  scale_color_gradient(name = "Saltwater Vulnerability",
                       low = "bisque", high = "red") +
  theme(legend.position = "bottom") +
  labs(title = "Saltwater Vulnerability") +
  theme(text = element_text(family = "Times New Roman"))

```

# Loop through all variables once the above is done. 

```{r}
variables_to_loop <- names(plot_data)[9:length(names(plot_data))]

plot_list <- list()

for (variable in variables_to_loop) {
  plot <- ggplot(data = plot_data) +
    geom_sf(color = "blue", size = 1) +  
    geom_sf(aes_string(color = variable)) +  
    scale_color_continuous(name = variable) +  
    scale_size_continuous(name = "",
                          labels = scales::comma_format(),
                          range = c(0.1, 10)) +  
    labs(title = variable)  
  
  plot_list[[variable]] <- plot
}


print(plot_list[[1]])

for (i in seq_along(plot_list)) {
  ggsave(paste0("plot_", names(plot_list)[i], ".png"), plot_list[[i]], width = 8, height = 6)
}

```