---
title: "tests_for_map_points"
output: html_document
date: "2024-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}
library(pacman)
library(ipumsr)
library(plyr)
library(dplyr)
library(readr)
library(sf)
library(xml2)
library(terra)
library(usmap)
p_load(tidycensus,tidyverse,sf,data.table,tigris,
       srvyr,survey,rio,car,janitor,rmapshaper,scales,santoku,viridis, urbnmapr, usmap)
```


## Current map_points function
See if reach codes are necessary or if polygons can be converted to longitude and latitude points 
```{r}
# countryonevariable %>%
#   dplyr::mutate(lon = sf::st_coordinates(.)[,2],
#                 lat = sf::st_coordinates(.)[,1])
```


```{r}
# map_points <- function(shapefile, data, location_columns, variable_name, base_unit) {
#   # Args: 
#     #shapefile: optional file path providing relevant boundaries e.g state, county, HUC. If no file to provide, write NULL 
#     #data: dataset containing points to be visualized on the map 
#     #location_columns: a list of column names (strings) that contain location data e.g c('lat', 'lon') or c('State')
#     #variable_name: a list of column names (string) that contain the variables of interest to be mapped e.g c('saltwater', 'Population Total')
#   # Outputs; 
#     #Points on a map of the U.S. Will produce a map for each variable name listed in variable_name
#   require(usmap)
#   
#   for (variable in variable_name) {
#     if (length(location_columns) > 0) {
#       
#       # Read and convert files to shapefiles 
#       if (!is.logical(shapefile) || !is.null(shapefile)) {
#         if (location_columns[1] %in% colnames(variable_name)) {
#         #check if location_columns >1, if more, error 
#         shapefile <- st_make_valid(shapefile)
#         # Ensure that geometries are points, lines, or polygons
#         if (any(st_geometry_type(shapefile) %in% c("POINT", "MULTIPOINT", "LINESTRING", "MULTILINESTRING", "POLYGON", "MULTIPOLYGON"))) {
#           if (!inherits(map, "sf")) {
#             stop("The object is not an sf object.")
#             }
#           shapefile <- shapefile %>%
#             mutate(centroid = st_centroid(geometry)) %>%
#             mutate(lon = st_coordinates(centroid)[, 1],
#                    lat = st_coordinates(centroid)[, 2])
#           map_data <- data %>%
#             select(location_columns[1], location_columns[2], variable) %>%
#             usmap_transform()
#         
#         # Plot combined data
#         p <- usmap::plot_usmap() +
#         geom_sf(data = map_data, color = "red") +
#         theme_minimal() +
#         labs(title = "Mapping the United States",
#              subtitle = variable,
#              x = "Longitude",
#              y = "Latitude")
#       print(p)
#       } else {
#             stop("Unsupported geometry type. Please ensure the shapefile contains points, lines, or polygons.")
#       }
#         } else {
#         map <- st_as_sf(shapefile)
#         usmap <- usmap_transform(map)
# 
#         plot_data <- usmap %>% left_join(data, by = location_columns, relationship = "many-to-many")
# 
# 
#         p <- plot_usmap(base_unit, color = "gray80") +
#           geom_sf(data = plot_data, color = "lightblue", size = 0.2) +
#           geom_point(data = plot_data,
#                        aes(geometry = geometry),
#                      alpha = 0.2,color = "red",
#                      stat = "sf_coordinates") +
#           scale_size_continuous(name = variable, labels = scales::comma_format(), range = c(0.01, 7)) +
#           scale_color_gradient(name = variable, low = "bisque", high = "red") +
#           theme(legend.position = "bottom") +
#           labs(title = variable) +
#           theme(text = element_text(family = "Times New Roman"))
#         print(p)
#         }
#       }
#         } else {
#     
#       map_data <- data %>%
#         select(location_columns[1], location_columns[2], variable) %>%
#         #NOT GENERAL 
#         rename(lat = LATITUDE, lon = LONGITUDE) %>%
#         usmap_transform()
#       
#       
#       # Plot combined data
#       p <- usmap::plot_usmap() +
#         geom_sf(data = map_data, color = "red") +
#         theme_minimal() +
#         labs(title = "Mapping the United States",
#              subtitle = variable,
#              x = "Longitude",
#              y = "Latitude")
#       print(p)
#       }
#       } else {
#       print('Please enter a location column name')
#     }
#   }
# }

```

```{r}
# map_points(map, data, c('lon', 'lat'), 'SWIVULN', 'counties')

```



```{r}
# #making a simple subfunction based on saltwater intrusion case
# join_map_points <- function (shapefile, data, location_columns, variable_name, base_unit) {
#   #processing shapefile 
#   shapefile <- st_as_sf(shapefile)
#   shapefile <- st_make_valid(shapefile)
#   usmap <- usmap_transform(shapefile)
#   
#   plot_data <- usmap %>% left_join(data, by = location_columns[1], relationship = "many-to-many")
#   
#   plot_usmap(base_unit, color = "gray80") + 
#   geom_sf(data = plot_data, color = "lightblue", size = 0.2) +
#   geom_point(data = plot_data,
#     aes(color = SWIVULN, size = SWIVULN, geometry = geometry),
#     alpha = 0.2,
#     color = "red",
#     stat = "sf_coordinates"
#   ) +
#   scale_size_continuous(name = "Saltwater Vulnerability",
#                         labels = scales::comma_format(),
#                         range = c(0.01, 7)) + 
#   scale_color_gradient(name = "Saltwater Vulnerability",
#                        low = "bisque", high = "red") +
#   theme(legend.position = "bottom") +
#   labs(title = "Saltwater Vulnerability") +
#   theme(text = element_text(family = "Times New Roman"))
# 
#   
# }

```

```{r}
# join_map_points(map, data, c("REACHCODE"), NULL, 'counties')
```
## Second Iteration of the function 

```{r}
# There are 3 cases for inputs: 
# 1: Shapefile with a location label (str) AND a dataset with a location label (str). 
# 2: Dataset with lat, lon columns or a geometry containing lat. 
# 3: Shapefile that can automatically be mapped with usmap. Please check the package's location data requirements for this. 

map_points <- function(shapefile, data, location_columns, variable_name, base_unit) {
  
  # Args: 
    #shapefile: optional file path providing relevant boundaries e.g state, county, HUC. If no file to provide, write NULL 
    #data: dataset containing points to be visualized on the map 
    #location_columns: a list of column names (strings) that contain location data e.g c('lat', 'lon') or c('State')
    #variable_name: a list of column names (string) that contain the variables of interest to be mapped e.g c('saltwater', 'Population Total')
    #base_unit: a string with the base unit to be drawn on the usmap e.g 'counties'
  # Outputs; 
    #Points on a map of the U.S. Will produce a map for each variable name listed in variable_name
  
  #For case type 1 
  if (length(location_columns) == 1) {
  
    usmap <- usmap_transform(shapefile)
   
    plot_data <- usmap %>% left_join(data, by = location_columns[1], relationship = "many-to-many")
    
    p <- plot_usmap(base_unit, color = "gray80") + 
      geom_sf(data = plot_data, color = "lightblue", size = 0.2) +
      geom_point(data = plot_data,
                 aes(color = NULL, size = NULL, geometry = geometry),
                 alpha = 0.2,
                 color = "red",
                 stat = "sf_coordinates") +
      scale_size_continuous(name = "Saltwater Vulnerability",
                            labels = scales::comma_format(),
                            range = c(0.01, 7)) +
      scale_color_gradient(name = "Saltwater Vulnerability",
                           low = "bisque", high = "red") +
      theme(legend.position = "bottom") +
      labs(title = "Saltwater Vulnerability") +
      theme(text = element_text(family = "Times New Roman"))
    print(p)

    #For case type 2
  } else if (length(location_columns) == 2) {
      map_data <- data %>%
        select(location_columns[1], location_columns[2], variable_name) %>%
        rename(lat = location_columns[1], lon = location_columns[2]) %>%
        usmap_transform()
        
        # Plot combined data
      p <- usmap::plot_usmap() +
        geom_sf(data = map_data, color = "red") +
        theme_minimal() +
        labs(title = "Mapping the United States",
             subtitle = variable_name,
             x = "Longitude",
             y = "Latitude")
      print(p)
    
    #For case type 3
    } else if (is.null(location_columns)) {
      # map_data <- data %>%
      #   select(location_columns[1], location_columns[2], variable) %>%
      #   rename(lat = LATITUDE, lon = LONGITUDE) %>%
      #   usmap_transform()
    
      # Plot combined data
      shapefile <- shapefile %>%
        rename(lat = location_columns[1], lon = location_columns[2]) %>%
        usmap_transform()
      
      p <- usmap::plot_usmap() +
        geom_sf(data = shapefile, color = "red") +
        theme_minimal() +
        labs(title = "Mapping the United States",
             subtitle = variable_name,
             x = "Longitude",
             y = "Latitude")
      print(p)
    }
  }

```
## Saltwater Intrusion Test Case

```{r}
#pre-processing for saltwater intrusion
data <- read_csv("~/Downloads/SGD_Coastal_Vulnerabilities.csv", 
                 col_types = cols(.default = "c"))


map <- vect("~/Downloads/Catchment_CONUS_coastline/Catchment_CONUS_coastline.shp")

data$REACHCODE <- ifelse(nchar(data$REACHCODE) < 14, 
                          paste0("0", data$REACHCODE), 
                          data$REACHCODE)

data <- data %>%
  group_by(REACHCODE, REGION) %>%
  summarize(SWIVULN = sum(as.numeric(SWIVULN)))

# head(data)
# head(map)
```


```{r}
map <- st_as_sf(map)
```

```{r}
map_points(map, data, c("REACHCODE"), c('SWIVULN'), 'counties')
```

## CWNS Test Case

Testing on cwns point data with latitude, longitude, and measure of interest (CURRENT_DESIGN_FLOW)
This works on Nora's computer but it includes the islands, so the projection is a little funky. 
Would be helpful to standardize the projection and cut out extraneous islands from the mapping 

```{r}
cwns <- read_csv("~/Downloads/cwns-flow.csv")
```

```{r}
# colnames(cwns)
map_points(NULL, cwns, c('LATITUDE', 'LONGITUDE'), 'CURRENT_DESIGN_FLOW', NULL)
```

## Interbasin Transfers Test Case
```{r}
ibt <- read_sf(dsn = '~/Downloads/IBT Geospatial Data/', layer = "Digitized_IBTs")
```


```{r}
map_points(ibt, NULL, NULL, NULL, NULL)
```

## Helper function: pull lat/lon from geometry
```{r}
# #convert geometry to lat and lon 
# #could be a helper function if the case comes up in a dataset I haven't tested 
# if (!is.logical(shapefile) || !is.null(shapefile)) {
#       shapefile <- st_make_valid(shapefile)
#       # Ensure that geometries are points, lines, or polygons
#       if (any(st_geometry_type(shapefile) %in% c("POINT", "MULTIPOINT", "LINESTRING", "MULTILINESTRING", "POLYGON", "MULTIPOLYGON"))) {
#         if (!inherits(map, "sf")) {
#           stop("The object is not an sf object.")
#         }
#         shapefile <- shapefile %>%
#         mutate(centroid = st_centroid(geometry)) %>%
#         mutate(lon = st_coordinates(centroid)[, 1],
#                lat = st_coordinates(centroid)[, 2])
#       }
#       }
#   
```
