---
title: "plotting_layers"
author: "Mbye Sallah"
date: "2024-07-23"
output: html_document
---

#### Import libraries
```{r}
library(tidyverse)
library(leaflet)
library(leaflet.providers)
library(sf)
library(rjson)
library(htmltools)
library(geojsonio)
```

# See all the functions in leaflet
```{r}
# ls("package:leaflet")
```

#### Create a map widget
```{r}
leaflet_map <- leaflet() %>% 
  setView(-96, 37.8, zoom = 4) %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addTiles(urlTemplate = "https://tiles.wmflabs.org/hikebike/{z}/{x}/{y}.png")
  leaflet_map
```


#### read the GeoJSON file
```{r}
counties_sf <- st_read("/Users/mbye/DSSG/WaterReuseDSSG2024/geo-ndxr/prototype-dev/test_mapping/test/counties.geojson")
```


#### Plot geojson shapefile on map
```{r}
leaflet_shapes_map <-  leaflet_map %>% 
  addTiles() %>%
  addPolygons(data = counties_sf)
leaflet_shapes_map
```



#### Test mapping choropleth
##### Generate a new column in counties sf with values between 0 to 1
```{r}
counties_sf <-  counties_sf %>%
  mutate(
    index_value = runif(n()),
    index_value2 = runif(n(), min = 0, max = 100))
```

##### Save data to geoJSON
```{r}
geojson_write(counties_sf, file = "~/DSSG/WaterReuseDSSG2024/geo-ndxr/prototype-dev/test_mapping/counties_sf.geojson")
```


#### Create a color palette based on the index_value column
```{r}
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = counties_sf$index_value
)
```


#### Create the leaflet map and add polygons with the index_value column

```{r}
leaflet_shapes_map <- leaflet(counties_sf) %>% 
  addPolygons(
    fillColor = ~pal(index_value),
    fillOpacity = 0.7,
    color = NA,
    weight = 1,
    opacity = 1,
    highlight = highlightOptions(
      weight = 3,
      color = NA,
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste("Index Value:", round(index_value, 2))
  ) %>% 
  addLegend(
    pal = pal,
    values = ~index_value,
    title = "Index Value",
    opacity = 0.7,
    position = "bottomright"
  ) %>% setView(-96, 37.8, zoom = 4)


# Display the map
leaflet_shapes_map
```


#### Save the map in html
```{r}
save_html(leaflet_shapes_map,
           file = "/Users/mbye/DSSG/WaterReuseDSSG2024/geo-ndxr/prototype-dev/test_mapping/index.html")
```


#### Do the same thing using the web_mapper function
```{r}
source("~/DSSG/WaterReuseDSSG2024/geo-ndxr/prototype-dev/web_mapper.R")
```


#### Map index data
```{r}
random_index <- web_mapper(data_with_shapes = counties_sf,
           index_value_column = "index_value",
           name_column = "NAME")
random_index
```


