---
title: "process_data"
output: html_document
date: "2024-08-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Index and Website Creation Pipeline 

Creator pipeline for calculating a default index and setting up an interactive website displaying an index map. This is a working template for index creation. DO NOT PUSH to the GitHub repository. NOTE: set gitignore? for this file 

## Required Packages and source GeoNdxR crosswalk functions 
Ensure that the config.yml file in your downloaded GitHub repo contains your correct home path or set your working directory to the highest level of the GitHub repo. 

```{r}
#Require / Load libraries

require(tigris, tidyverse, dplyr, sf, usmap, ggplot2, naniar, raster, terra, stars, nngeo, exactextractr, naniar, geojsonsf, rmapshaper, geojsonio)

#Source GeoNdxR crosswalk functions. Find documentation for these functions linked in the GitHub README. 
source("/code/crosswalk/crosswalk_spatial.R")
source("/code/crosswalk/crosswalk_census.R")
source("/code/crosswalk/crosswalk_id.R")
source("/code/crosswalk/apply_weights.R")
source("/code/website/helper/edit_website.R")
source("/code/create_index/driver_weighted_pca.R")

```


## Load, explore, and transform data 
Users should provide geospatial dataset(s) to be included in the index. Be mindful of filetypes and requirements. Code chunk below commented with further instruction. 

```{r}

data_path <- "~/Downloads/data"

#If you are loading a shapefile (as opposed to csv, geojson, etc.), include folder with all relevant files in the file path

file_name <- "shape_file_folder/example.shp"

#Example of loading a shapefile. Adjust read/load R function based on file type. 
dataset1 <- #read_sf(paste0(data_path, file_name))

## Repeat for all data included in the index. If you only have one file, ignore the following commented out code.
# 
# file_name2 <- "example.csv"
dataset2 <- #read.csv(paste0(data_path, file_name))

```

Add chunks to explore data. Create plots, use the head(data) function, inspect NAs with naniar function gg_miss_var(data). 

## Crosswalk data to a target scale

```{r}
#Use crosswalk functions to match datasets to a target scale. 
#Please use the crosswalk tutorial file located in the docs folder of the GitHub repo for more guidance. 
#Convert all datasets to be sf or st objects that share the same CRS. From the sf package, use st_read and st_transform respectively to achieve this. 

##Example crosswalk using crosswalk_spatial function 
crosswalk_data <- #crosswalk_spatial(dataset1, dataset2, extensive = TRUE, join_method = "areal_weighted")

```

## Convert data to percentiles
```{r}
##Data must be numeric (except ID column) and converted to percentiles to be included in the index. 
##Data must contain only one geospatial ID column (ex: "County Name", or "PWSID"). Data must contain a geometry column. All other columns must be numeric. 

percentile_data <- #calculate_percentiles(crosswalk_data, "id_column")

```

